\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
{\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

Throughout the writing of this thesis, I have received a great deal of support and assistance. I would first like to thank the company Motius for supporting me in the evaluation and providing me the dataset.

I want to acknowledge my supervisor, Professor Diepold, for accepting my thesis topic.

I would also like to thank Julian Wörmann, who guided me, brainstormed with me, showed interest, and gave feedback about my thesis. I wish him the best with his career as a doctor and also on his new adventure with his family.

Besides, I would like to thank my parents, Zübeyde and Necmi Pekmezci. I would not be here without your emotional and financial support from the day that I was born.

I want to thank my girlfriend, Cosku Özdemirci, who reminded me how to feel emotions and made life more bearable.

I also want to thank my Munich friends for making life more enjoyable, especially Berkay Soykan, Haydar Sahin, Helin Senbayram, Ulas Özdemirci, Zeynep Sasi and Zeynep Karagoz, who gave me the idea of using this section more productively.

My long-distance friends Ada Gencoglu and Ugur Kocak also supported me emotionally, and I am thankful for having them in my life.

I want to offer special thanks to Akin Akkan, who, although no longer with us, our memories will be in my mind until I die.

I also want to thank some people that I do not know personally: thanks to the organizers of the Fusion Festival for teaching how to resist against the authority when needed and I want to thank the elected mayor of Istanbul, Ekrem Imamoglu, for giving me motivation, hope and showing me how to win hearts without being exclusive or hostile.


\cleardoublepage{}
